### 2015-06-14 writing/S'15 LSTM/intro.tex
Paper push.

Neural network models that encode sentences as vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}. Many of the most successful of these models are based on either tree models or sequence models. sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical sturcture of linguistic phrases.

Tree models are presented as the more principled choice, since they directly exploit the principle of compositionality---a fundamental idea in linguistic semantics \cite{Partee84,Janssen97}---which states that the meanings of phrases and sentences are constructed recursively from the meanings of their parts in a specific syntactically-guided order.

While both classes of model can perform objectively well on many tasks, and both are under active development, tree models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}. 
In this paper, we use focused artificial data experiments to expose a possible explanation for these results: standard sequence models can learn to implicitly exploit recursive compositional structure.

@@

As in \newcite{Bowman:Potts:Manning:2014}, we find that standard tree neural networks are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size. We also find that extending the training set to include larger structures mitigates this decay. In addition, we find that a sequence model centered on a single-layer LSTM is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree models.

>>>>

Neural network models that encode sentences as vectors have been successfully used in a wide array of NLP tasks, including translation \cite{sutskever2014sequence}, parsing \cite{dyer2015transition}, and sentiment analysis \cite{tai2015improved}. Many of the most successful of these models are based on either tree models or sequence models. sequence models based on recurrent neural networks build representations incrementally from left to right \cite{elman1990finding,sutskever2014sequence}, while tree models based on \ii{recursive} neural networks \cite{goller1996learning,socher2011semi} build representations incrementally according the hierarchical sturcture of linguistic phrases. Tree models are presented as the more principled choice, since they align with standard linguistic assumptions about constituent structure and the compositional derivation of complex meanings.

While both classes of model can perform objectively well on many tasks, and both are under active development, tree models have not shown the kinds of dramatic performance improvements over sequence models that their billing would lead one to expect: head-to-head comparisons with sequence models show either modest improvements \cite{tai2015improved} or none at all \cite{li2015tree}. 
In this paper, we use focused artificial data experiments to expose a possible explanation for these results: when trained on strings that contain indications of recursive syntactic structure, standard sequence models implicitly learn tree-like constituent structures even though they do not process their inputs as trees. That is, to the extent that constituent structure it useful, they can learn approximations of it without being required to do so by their architecture.

@@

As in Bowman et al., we find that standard tree neural networks are able to learn the necessary generalization, with their performance decaying gradually as the structures in the test set grow in size. We also find that extending the training set to include larger structures mitigates this decay. In addition, we find that a sequence model centered on a single-layer LSTM is also able to generalize to unseen large structures, but that it only does this when trained on a larger and more complex training set than is needed by the tree models.
